---
title: "Ratio Estimator"
author: 
  - Shreya Sakura Noskor
  - Zeyi Cai
  - Mo (Molly) Zhou
  - Bingxu Li
thanks: "Code and data are available at: [https://github.com/NotSakura/RatioEstimator.git](https://github.com/NotSakura/RatioEstimator.git)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(haven)
```

```{r}
#| include: false
#| warning: false
#| message: false


ipums_extract <- read_dta("../data/01-raw_data/usa_00001.dta")
```


# Instructions on Extracting Data

We first go to the IPUMS american data website and then we click 'get data' [https://usa.ipums.org/usa/](https://usa.ipums.org/usa/). Next we click SELECT SAMPLES then we deselect everything there. We then click ASC 2022 to get the 2022 data. Then we check which values we want so we go to Households and pick Geographic, where we select STATEICP. And we go to Person and click education where we click EDUC. Once we finished that we click view cart to see our dataset (click CREATE DATA EXTRACT). We change the data format to .dta and we change the sample size to 500 after we click CUSTOMISE SAMPLE SIZES. Then we give and an appropriate detail fo extract that includes today's date and what columns we are going to be looking at. Then we download it (after making an account) and add it to our repository to read. 



```{r}
#| include: false
#| warning: false
#| message: false
library(dplyr)

# Assume ipums_extract is your original dataset
# Step 1: Filter for California (State Code 71)
california_data <- ipums_extract %>% filter(stateicp == 71)

# Step 2: Total number of respondents in California
total_respondents_california <- nrow(california_data)

# Step 3: Number of respondents with doctoral degrees in California
doctoral_respondents_california <- california_data %>%
  filter(educd == 116) %>%  # Ensure 116 corresponds to doctoral degrees
  nrow()

# Step 4: Calculate the ratio for California
ratio_california <- doctoral_respondents_california / total_respondents_california

# Step 5: Estimate total respondents for each state using ipums_extract dataset
estimated_data <- ipums_extract %>%
  group_by(stateicp) %>%
  summarize(
    doctoral_respondents = sum(educd == 116, na.rm = TRUE),  # Count of doctoral respondents
    actual_respondents = n()                                   # Total respondents
  ) %>%
  mutate(
    estimated_total = (doctoral_respondents / ratio_california)   # Apply California's ratio
  )



```

# brief overview of the ratio estimators approach

Refered to the textbook to follow the approch of finding the ratio estimators. we first filtered for the state of california. Then we counted how many actual respondants there were and how many of them have their doctorate. We then divide it to find the ratio and then we apply that to the number of doctorate person in each state. 


# Ratio and Actual Number of Respondants

```{r}
#| echo: false
#| warning: false
#| message: false

library(knitr)


estimated_data |>
  kable(
    col.names = c("State ICP", "Doctoral Respondents", "Actual Respondents", "Estimated Total Respondents"),
    format.args = list(big.mark = ","),
    digits = 2
  )
```



# Explanation for why they are different

California is home to many  universities and research institutions, which may contribute to a higher number of doctoral degree holders relative to the population. This uniqueness can skew the ratio when applied to other states that do not share the same educational environment.

Also the ratio estimator assumes that the relationship between doctoral respondents and total respondents in California is same for all other states. This may not be true as in some states there may not be as much universities or the proportion of regular students to doctorate students could be different. 

\newpage


# References


